---
title: Workflow for Creating a New Mapping Project
format:
  html:
    code-fold: true
    css: styles.css
execute:
  warning: false
---


```{r}
#| message: false
library(sf)
library(tmap)
tmap_mode("plot")
project_gpkg <- "sample_project/demo.gpkg"
disturbance_gpkg <- "G:/Shared drives/Beacons data/Public/geopackage_creator/bp_datasets.gpkg"
aoi <- st_read(project_gpkg, layer = "studyarea", quiet = TRUE)
aoi_bbox <- st_as_text(st_as_sfc(st_bbox(st_buffer(aoi, 50000))))
bnd <- st_read(disturbance_gpkg, layer = "bnd", quiet = TRUE)
grid <- st_read(disturbance_gpkg, layer = "grid", quiet = TRUE)
line <- st_read(disturbance_gpkg, layer = "linear_disturbances", wkt_filter = aoi_bbox, quiet = TRUE)
poly <- st_read(disturbance_gpkg, layer = "areal_disturbances", wkt_filter = aoi_bbox, quiet = TRUE)
map_bbox <- st_bbox(st_buffer(aoi, 50000))
```

## Introduction

This workflow should be used when creating a new project and merging it with the most recent disturbance database. Steps include:

1. Prepare project geopackage and backup the existing disturbance database
2. Use the new project AOI to select all intersecting 10-km grid cells
3. Use the selected grid cells to clip all existing linear and areal features
4. Digitize new features, remove features no longer visible, and modify erroneous features
5. Erase all linear and areal features within the AOI-GRID from the existing disturbance database
6. Replace AOI-GRID gap with the newly modified linear and areal disturbance features
7. Replace the original disturbance database with the updated version

## Step 1: Backup existing disturbance database

Prepare the project geopackage by creating a new project folder and copying your AOI map layer into it. Also, make sure to backup or archive the existing disturbance database (disturbances.gpkg) e.g., `disturbances_2026-08-13.gpkg`

```{r}
#| fig-width: 11
#| fig-height: 11
tm_shape(bnd, bbox = map_bbox) + tm_borders(col = "black") +
  tm_shape(poly) + tm_polygons(fill = "blue", col = NA) +
  tm_shape(line) + tm_lines(col = "blue") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("Area of interest")
```

## Step 2: Select intersecting grid cells

Use the AOI to select all intersecting 10x10 km grid cells.

```{r}
#| fig-width: 11
#| fig-height: 11
aoi_grid <- grid[st_intersects(grid, aoi, sparse = FALSE)[, 1], ]
st_write(aoi_grid, project_gpkg, layer = "aoi_grid", delete_layer = TRUE, quiet = TRUE)
tm_shape(grid, bbox = map_bbox) + tm_borders(col = "grey50") +
  tm_shape(aoi_grid) + tm_polygons(fill = "grey70", col = "grey50") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("Selected AO grid cells")
```

## Step 3: Clip existing linear and areal features

Clip existing linear and areal features to the aoi_grid extent and save them to the project geopackage.

```{r}
#| fig-width: 11
#| fig-height: 11
aoi_grid_line <- st_intersection(line, st_union(aoi_grid))
aoi_grid_poly <- st_intersection(poly, st_union(aoi_grid))
#st_write(aoi_grid_line, project_gpkg, layer = "aoi_grid_line", delete_layer = TRUE, quiet = TRUE)
#st_write(aoi_grid_poly, project_gpkg, layer = "aoi_grid_poly", delete_layer = TRUE, quiet = TRUE)
tm_shape(aoi_grid, bbox = map_bbox) + tm_borders(col = "grey50") +
  tm_shape(aoi_grid_poly) + tm_polygons(fill = "blue", col = NA) +
  tm_shape(aoi_grid_line) + tm_lines(col = "blue") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("Clipped disturbance features")
```

## Step 4: Digitize

Use QGIS to digitize new features, remove features no longer visible, and modify erroneous features. Make sure the entire area of the `aoi_grid` is reviewed and updated.

```{r}
#| fig-width: 11
#| fig-height: 11
aoi_grid_line <- st_intersection(line, st_union(aoi_grid))
aoi_grid_poly <- st_intersection(poly, st_union(aoi_grid))
#st_write(aoi_grid_line, project_gpkg, layer = "aoi_grid_line", delete_layer = TRUE, quiet = TRUE)
#st_write(aoi_grid_poly, project_gpkg, layer = "aoi_grid_poly", delete_layer = TRUE, quiet = TRUE)
tm_shape(aoi_grid, bbox = map_bbox) + tm_borders(col = "grey50") +
  tm_shape(aoi_grid_poly) + tm_polygons(fill = "red", col = NA) +
  tm_shape(aoi_grid_line) + tm_lines(col = "red") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("New & revised disturbance features")
```

## Step 5: Erase existing features from the disturbance database

Erase all linear and areal features within the AOI-GRID from the existing disturbance database

```{r}
#| fig-width: 11
#| fig-height: 11
line2 <- st_difference(line, st_union(aoi_grid))
poly2 <- st_difference(poly, st_union(aoi_grid))
aoi_grid2 <- st_union(aoi_grid)
tm_shape(aoi_grid2, bbox = map_bbox) + tm_borders(col = "grey50") +
  tm_shape(poly2) + tm_polygons(fill = "blue", col = NA) +
  tm_shape(line2) + tm_lines(col = "blue") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("AOI-GRID gap erased from disturbance database")
```

## Step 6: Replace erased features with modified features

Replace AOI-GRID gap with the newly modified linear and areal disturbance features

```{r}
#| fig-width: 11
#| fig-height: 11
line2 <- st_difference(line, st_union(aoi_grid))
poly2 <- st_difference(poly, st_union(aoi_grid))
aoi_grid2 <- st_union(aoi_grid)
tm_shape(aoi_grid2, bbox = map_bbox) + tm_borders(col = "grey50") +
  tm_shape(poly2) + tm_polygons(fill = "blue", col = NA) +
  tm_shape(line2) + tm_lines(col = "blue") +
  tm_shape(aoi_grid_poly) + tm_polygons(fill = "red", col = NA) +
  tm_shape(aoi_grid_line) + tm_lines(col = "red") +
  tm_shape(aoi) + tm_borders(col = "black", lwd = 2) +
  tm_title("Updated disturbance database with AOI-GRID features")
```

## Step 7: Replace the original disturbance database with the updated version

Replace the original disturbance database with the updated version
